\textbf{Analysis of CSV File Reading in R}

### 2021-12-10 paper.Rmd
paper and slides updates

The goal of this analysis is to determine the effects of three factors - `R` package, file size, computer device - on the reading time for CSV files measured in seconds. Because of this, we chose to design our experiment with a  $3^3$ factorial experiment [@montgomery] in mind. Furthermore, we chose to include $n=20$ replicates for each combination of our factors.  Our chosen design implies that the total number of observations for our experiment is 540. Our data was generated by considering the $3^3$ possible combinations of `R` package, file size, and computer. We then measured the twenty replicates for each combination according to the order of a random sample of 1 through $3^3$.

@@

The response variable, $y_{ijkl}$, indicates the speed time in seconds corresponding to each run conducted in the experiment. Here $\mu$ represents the overall mean reading time,  and $\tau_i$, $\beta_j$, and $\gamma_k$ are the main effects of three factors file size, computer device, and `R` function/package. There are also 3 two-factor and 1 three-factor interaction effects in our model. The assumptions for the model are that the residuals $\epsilon_{ijkl}$ are normally distributed and that the variance is constant across the groups.

@@

As a result of violations of normality and homoscedasticity, we then considered a couple of data transformation methods, namely, a log transformation and a Box-Cox power transformation [@boxcox] as possible ways to resolve the non-normality issue of our data. However, transforming the response variable, file reading time, did not improve the fit of this model and our assumptions are still not met. The ANOVA table, visual summary, and tests for normality and constant variance for these two methods can be found in the Appendix.

@@

One particularly interesting observation we had after conducting analysis of variance to our data is that  `readr::read_csv()` is the fastest method when running on the MacBook device, but on the other hand, the slowest for the other two computers.  Thus, we are interested in performing a quick simulation study to see if there is a function/package effect on the file reading time for the MacBook device only. In addition to the three data files with sizes 0.49, 1.22, 1.96 GB used in the previous experiment, we simulated seven additional datasets with sizes 1.51, 1.73, 2.16, 2.36, 2.61, 2.94, 3.32 GB

We then conducted an analysis of covariance (ANCOVA) with time elapsed as the response variable,  the file reading function as the categorical independent variable, and the file size as the continuous covariate.

The table output suggests that the mean reading time differs for all `R` CSV reading functions, after accounting for the effect of file size. Furthermore, both normality and constant variance conditions are met for this model, as illustrated by a Shapiro-Wilk test for normality and a Levene Test for homogeneity of variance.

We can then proceed to perform multiple comparisons to see which algorithms are different from each other. It is confirmed that the mean reading time for `readr::read_csv()` is different (faster) than the mean reading time of the other two techniques, and the other two also don’t differ in terms of average importing time.

@@

log transfomation

```{r}
mod <- lm(log(time) ~ computer * size * package, data = res)
mod %>% 
  anova() %>% 
  tidy() %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  knitr::kable(caption = "ANOVA for the three-factor factorial ANOVA model.")
```

@@

@@

Box-Cox Transformation -0.2543644

```{r}
mod <- lm(time ^ (-0.2543644) ~ computer * size * package, data = res)
mod %>% 
  anova() %>% 
  tidy() %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  knitr::kable(caption = "ANOVA for the three-factor factorial ANOVA model.")
```

@@

@@

>>>>

The goal of this analysis is to determine the effects of three factors - `R` package, file size, computer device - on the reading time for CSV files measured in seconds. Because of this, we chose to design our experiment with a  $3^3$ factorial experiment [@montgomery] in mind. Furthermore, we chose to include $n=20$ replicates for each combination of our factors.  Our chosen design implies that the total number of observations for our experiment is 540. Our data was generated by considering the $3^3$ possible combinations of `R` package, file size, and computer. We then obtained the twenty replicates for each combination according to the order of a random sample of 1 through $3^3$.

@@

The response variable, $y_{ijkl}$, indicates the speed time in seconds corresponding to each run conducted in the experiment. Here $\mu$ represents the overall mean reading time,  and $\alpha_i$, $\beta_j$, and $\gamma_k$ are the main effects of three factors file size, computer device, and `R` function/package. There are also 3 two-factor and 1 three-factor interaction effects in our model. In addition, this model includes an error component $\epsilon_{ijkl}$, and the assumptions for the residuals are that they are normally distributed and the variance is constant across the groups.

@@

As a result of violations of normality and homoscedasticity, we then considered a couple of data transformation methods, namely, a natural log transformation and a Box-Cox power transformation [@boxcox] as possible ways to resolve the non-normality issue of our data. However, transforming the response variable, file reading time, did not improve the fit of this model and our assumptions are still not met. The ANOVA table, visual summary, and tests for normality and constant variance for these two methods can be found in the Appendix.

@@

One particularly interesting observation we had after conducting analysis of variance to our data is that  `readr::read_csv()` is the fastest method when running on the MacBook device, but on the other hand, the slowest for the other two computers. This trend difference is depicted on the third plot of Figure 1. Thus, we are interested in performing a quick simulation study to see if there is a function/package effect on the file reading time for the MacBook device only. In addition to the three data files with sizes 0.49, 1.22, 1.96 GB used in the previous experiment, we simulated seven additional datasets with sizes 1.51, 1.73, 2.16, 2.36, 2.61, 2.94, 3.32 GB. After that we used each of the three methods (`read_csv()`, `import()`, `fread()`) to import the data files and record the time it took to read in. Figure 3 is a scatterplot of time elapsed and file size, broken down by CSV reading method, and it shows that `readr::read_csv()` is clearly the fastest data loading technique on the MacBook device.

```{r}
dat <- read_csv("add.csv") %>% 
  mutate(meth = factor(meth))
```

```{r, fig.height=3, fig.width=4, fig.cap="Scatterplot of time elapsed and file size, color coded by reading algorithm."}
dat %>% 
  ggplot(aes(size, elapsed, color = meth)) +
  geom_point() +
  geom_smooth(method = "lm", 
              size = 0.5,
              alpha = 0.3) +
  labs(x = "Size (GB)",
       y = "Time Elapsed (sec)",
       color = "Method") +
  scale_color_manual(values = c("midnightblue", "darkgreen", "darkorange2")) +
  theme(axis.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        axis.text = element_text(size = 8),
        legend.title = element_text(size = 9))
```

\vspace{-3mm}

We then conducted an analysis of covariance (ANCOVA) with time elapsed as the response variable, the file reading function as the categorical independent variable, and the file size as the continuous covariate. The ANCOVA model is given by \begin{equation} y_{ij} = \mu + \tau_i + \beta(x_{ij} - \bar{x}_{\cdot \cdot}) + \epsilon_{ij}, \end{equation}
where $\tau_i$ is the effect of the $i$th treatment, and $\beta_j$ is a parameter associated with the $j$th subject.

We were interested in using ANCOVA to investigate the effect of the packages/functions on the time elapsed. Table 5 suggests that the mean reading time differs among the CSV reading functions in `R`, after accounting for the effect of file size. Furthermore, both normality and constant variance conditions are met for this model, as illustrated by a Shapiro-Wilk test for normality and a Levene Test for homogeneity of variance (see Table 6).

```{r}
anc <- aov(elapsed ~ meth + size, data = dat)
a <- anc %>% 
  car::Anova(type = "III") %>% 
  tidy() %>% 
  mutate(across(where(is.numeric), ~ round(.x, 4)))

names(a) <- c("Term", "SS", "df", "F", "p-value")

a %>% 
  mutate(Term = str_replace(Term, "meth", "package")) %>% 
  knitr::kable(caption = "ANCOVA table for the analysis of reading time for different methods on MacBook device.")
```

```{r}
shap <- anc %>% 
  pluck("residuals") %>% 
  shapiro.test() %>% 
  tidy()
```

```{r}
lev <- leveneTest(elapsed ~ meth, data = dat) %>% 
  tidy() %>% 
  dplyr::select(-contains("df")) %>% 
  mutate(method = "Levene test for homogeneity of variance")
```

```{r}
shap %>% 
  bind_rows(lev) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  knitr::kable(caption = "Shapiro-Wilk and Levene tests for ANCOVA model.")
```

We can then proceed to perform multiple comparisons to see which algorithms are different from each other. It is confirmed by Table 7 that the mean reading time for `readr::read_csv()` is different (faster) than the mean reading time of the other two techniques, and the other two also don’t differ in terms of average importing time.

```{r}
library(multcomp)
glht(anc, linfct = mcp(meth = "Tukey")) %>% 
  tidy() %>% 
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>% 
  dplyr::select(-null.value, -term) %>% 
  rename(se = std.error,
         `adj p-value` = adj.p.value) %>% 
  knitr::kable(caption = "Pairwise comparisons with Tukey contrasts following ANCOVA.")
```

# Discussion

# Supplementary Material {-}

All materials related to this manuscript are publicly available on GitHub at \newline <https://github.com/qntkhvn/read_speed>.

@@

```{r}
mod <- lm(log(time) ~ computer * size * package, data = res)
mod %>% 
  anova() %>% 
  tidy() %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  knitr::kable(caption = "ANOVA table for the three-factor factorial ANOVA model with a natural log transformation to the response.")
```

\vspace{-2mm}

@@

```{r, fig.width=6, fig.height=3, fig.cap="Residuals plots for the three-factor factorial ANOVA model with a natural log transformation to the response."}
pqq + prvf
```

@@

\vspace{-4mm}

```{r}
shap %>% 
  bind_rows(lev) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3)),
         p.value = "0.000") %>% 
  knitr::kable(caption = "Shapiro-Wilk and Levene tests for the three-factor factorial ANOVA model with a natural log transformation to the response.")
```

```{r}
mod <- lm(time ^ (-0.2543644) ~ computer * size * package, data = res)
mod %>% 
  anova() %>% 
  tidy() %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  knitr::kable(caption = "ANOVA table for the three-factor factorial ANOVA model with a Box-Cox transformation to the response.")
```

\vspace{-2mm}

@@

```{r, fig.width=6, fig.height=3, fig.cap="Residuals plots for the three-factor factorial ANOVA model with a Box-Cox transformation to the response."}
pqq + prvf
```

@@

```{r}
shap %>% 
  bind_rows(lev) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3)),
         p.value = "0.000") %>% 
  knitr::kable(caption = "Shapiro-Wilk and Levene tests for the three-factor factorial ANOVA model with a Box-Cox transformation to the response.")
```

### 2021-12-10 slides.Rmd
paper and slides updates

---
title: Analysis of CSV File Reading in R
date: STAT 407 Fall 2021
output: 
  beamer_presentation: 
    theme: Boadilla
    color: beaver
    fonttheme: professionalfonts
    keep_tex: no
    df_print: kable
    highlight: tango
csl: apa.csl
bibliography: ref.bib
# aspectratio: 169
fontsize: 11pt
header-includes:
  - \usepackage{xcolor}
  - \definecolor{luc}{RGB}{88, 41, 49}
  - \setbeamercolor{itemize/enumerate body}{fg=luc}
  - \setbeamercolor{title}{fg=luc}
  - \setbeamercolor{frametitle}{fg=luc}
  - \setbeamertemplate{itemize item}{\color{luc}$\bullet$}
  - \setbeamertemplate{itemize subitem}{\color{luc}$\bullet$}
  - \setbeamercolor{footnote}{fg=luc}
  - \setbeamercolor{footnote mark}{fg=luc}
  - \setbeamercolor{footlinecolor}{fg=black,bg=lightgray}
  - \setbeamertemplate{navigation symbols}{}
  - \author[Q. Nguyen \& R. Tedesco]{Quang Nguyen* and Robert Tedesco*}
  - \institute[]{* \hspace{-1.5mm} Loyola University Chicago}
  - \setbeamercovered{transparent}
  - \setbeamercolor{block title}{fg=luc}
  - \setbeamercolor{local structure}{fg=luc}
  - \usepackage{bbding}
urlcolor: blue
---

@@

>>>>

---
title: \textbf{Analysis of CSV File Reading in R}
date: STAT 407 Fall 2021
output: 
  beamer_presentation: 
    theme: Boadilla
    color: beaver
    fonttheme: professionalfonts
    keep_tex: no
    df_print: kable
    highlight: tango
csl: apa.csl
bibliography: ref.bib
# aspectratio: 169
fontsize: 11pt
header-includes:
  - \usepackage{xcolor}
  - \definecolor{luc}{RGB}{88, 41, 49}
  - \setbeamercolor{itemize/enumerate body}{fg=luc}
  - \setbeamercolor{title}{fg=luc}
  - \setbeamercolor{frametitle}{fg=luc}
  - \setbeamertemplate{itemize item}{\color{luc}$\bullet$}
  - \setbeamertemplate{itemize subitem}{\color{luc}$\bullet$}
  - \setbeamercolor{footnote}{fg=luc}
  - \setbeamercolor{footnote mark}{fg=luc}
  - \setbeamercolor{footlinecolor}{fg=black,bg=lightgray}
  - \setbeamercolor{structure}{fg=luc}
  - \setbeamercolor{titlelike}{parent=structure}
  - \setbeamertemplate{navigation symbols}{}
  - \author[Q. Nguyen \& R. Tedesco]{Quang Nguyen* and Robert Tedesco*}
  - \institute[]{* \hspace{-1.5mm} Loyola University Chicago}
  - \setbeamercovered{transparent}
  - \setbeamercolor{block title}{fg=luc}
  - \setbeamercolor{local structure}{fg=luc}
  - \usepackage{bbding}
urlcolor: blue
---

@@

*   Generate 3 files of different sizes consisting of columns of randomly-draw samples from $N(0,1)$, then export to CSV.

*   $3^3$ possible combinations of `R` package, file size, and computer.

*   Use the 3 reading functions to load the CSV files, then `Sys.time()` to measure time elapsed.

*   Obtain 20 replicates for each combination according to the order of a random sample of 1 through $3^3$.
